Delocalization of Eigenvectors in Random Matrices

### 2025-03-20 Draft.tex
check

\subsection{Stopping time}
While the three terms we achieved in the previous section are a good start, we have non-linear powers w.r.t to term $\E_t(z)$ itself, which is what we are trying to bound. This means that we have to be more careful in our analysis. For this purpose, we will have to use a stopping time argument \cite{bandSDE}. More specifically, for a fixed $\delta_\sto>0$ and $D\lesssim 1$ ($a\lesssim b\equiv a = O(b))$, independent of $N$, let us define the following:
\begin{equation*}
\tau_{\text{stop}, 1}=\inf \left\{s\geq 0: \max_{a, b} |\E_s(z)_{ab}|\geq W^{\delta_{\sto}}W^{-\frac{3}{4}}\cdot W^{-1}|\im w_s|^{-1}|\im w_s|^{-\frac{1}{2}}\right\}\wedge 1
\tag{3.2.1}
\end{equation*}

@@

\section{Quantum Diffusion and Delocalization}
\label{sec:deloc}

>>>>

\subsection{Stopping time}
While the three terms we achieved in the previous section are a good start, we have non-linear powers w.r.t to term $\E_t(z)$ itself, which is what we are trying to bound. This means that we have to be more careful in our analysis. For this purpose, we will have to use a stopping time argument \cite{bandSDE}. More specifically, for a fixed $\delta_\sto>0$ and $D\lesssim 1$ ($a\lesssim b\equiv a = O(b))$, independent of $N$, let us define the following:
\begin{equation*}
\tau_{\text{stop}, 1}=\inf \left\{s\geq 0: \max_{a, b} |\E_s(z)_{ab}|\geq W^{\delta_{\sto}}W^{-\frac{3}{4}}|\im w_s|^{-1}\cdot W^{-1}|\im w_s|^{-\frac{1}{2}}\right\}\wedge 1
\tag{3.2.1}
\end{equation*}

@@

\noindent The intuition behind the latter stopping times is simple - $\tau_{\sto, 1}$ reflects the first occurrence of

The stopping times in equations (3.3)-(3.5) establish a bootstrapping mechanism that's fundamental to the proof. Let's analyze each component:

For $\tau_{\text{stop},1}$ in equation (3.3), we have:
\begin{align}
\tau_{\text{stop},1} := \inf\left\{s \geq 0 : \max_{a,b}|E_s(z)_{ab}| \geq W^{\delta_{\text{stop}}}W^{-\frac{3}{4}}|\text{Im}w_s|^{-1} \cdot W^{-1}|\text{Im}w_s|^{-\frac{1}{2}}\right\} \wedge 1
\end{align}

This controls the error term $E_s(z) = T_s(z) - \Theta_s$ between the actual flow $T_s(z)$ and the diffusion profile $\Theta_s$. The scaling factor $W^{-\frac{3}{4}}|\text{Im}w_s|^{-1} \cdot W^{-1}|\text{Im}w_s|^{-\frac{1}{2}}$ is carefully chosen for two reasons:

\begin{enumerate}
\item The typical size of $\max_{a,b}(\Theta_t)_{ab}$ is $W^{-1}|\text{Im}w_t|^{-\frac{1}{2}}$, which establishes the natural scale for comparison.

\item The additional factor $W^{-\frac{3}{4}}|\text{Im}w_s|^{-1}$ provides the necessary smallness to control nonlinear terms in the SDE flow for $E_s(z)$ (as seen in Lemma 7).
\end{enumerate}

The assumption $\eta \gg W^{-\frac{3}{4}}$ is critical because $\text{Im}w_t = \eta + (1-t)\text{Im}m(z)$ decreases monotonically from $\text{Im}w_0$ to $\text{Im}w_1 = \eta$. If $\eta$ were smaller than $W^{-\frac{3}{4}}$, then $|\text{Im}w_1|^{-1}$ would be too large to maintain control of $E_s(z)$, causing the error estimates to break down.

For $\tau_{\text{stop},2}$ in equation (3.4), we have:
\begin{align}
\tau_{\text{stop},2} := \inf\left\{s \geq 0 : \max_{a,b}\frac{|G_s(z)_{ab} - \delta_{ab}m(z)|^2}{(S^{1/2}T_s(z)S^{1/2})_{ab} + S^{1/2}_{ab} + W^{-D}} \geq W^{\delta_{\text{stop}}/10}\right\} \wedge 1
\end{align}

This controls the deviation of the resolvent entries from the semicircle law, normalized by appropriate terms. The denominator provides the natural scale for these deviations, with the addition of $W^{-D}$ ensuring regularity.

The self-reinforcing argument works as follows:

\begin{enumerate}
\item Initially, we know $E_0(z) = 0$, so $\tau_{\text{stop}} > 0$ with probability 1.

\item Assuming $s \leq \tau_{\text{stop}}$, we can control the error terms in the SDEs for $E_s(z)$ and $G_s(z)$.

\item These controlled estimates allow us to prove that $\mathbb{P}[\tau_{\text{stop}} < 1] \leq N^{-D}$ for any $D > 0$ (Theorem 8).

\item Therefore, with high probability, the error bounds hold for all $s \in [0,1]$, validating our assumptions.
\end{enumerate}

This is fundamentally a continuity argument: we assume the bounds hold up to $\tau_{\text{stop}}$, use them to prove $\tau_{\text{stop}} = 1$ with high probability, which in turn validates that the bounds hold throughout $[0,1]$.

The critical relationship between $W$ and $N$ (namely $W \gg N^{8/11}$) emerges precisely from ensuring that the error term $W^{-\frac{7}{4}}\eta^{-\frac{3}{2}}$ in equation (2.1) is much smaller than the typical size of $\Theta$ entries ($\approx W^{-1}\eta^{-\frac{1}{2}}$), while maintaining $\eta \approx W^2N^{-2}$, which corresponds to the diffusion time scale.

To further elaborate on the stopping time mechanism, we examine the error control in more detail:

\begin{enumerate}
\item The SDE for $E_t(z)$ in Lemma 7 contains nonlinear terms of the form $E_t(z)^2$. These terms require the smallness factor $W^{-\frac{3}{4}}|\text{Im}w_t|^{-1}$ to remain controlled.

\item When we solve the integral equation for $E_t(z)$, we obtain terms of the form $E_t^M(z)$, $E_t^D(z)$, and $E_t^S(z)$ (from the martingale part, the drift part, and the nonlinear self-interaction part, respectively).

\item The estimates in Lemmas 9, 10, and Proposition 11 show that each of these terms remains smaller than $W^{\delta_{\text{stop}}}W^{-\frac{3}{4}}|\text{Im}w_t|^{-1} \cdot W^{-1}|\text{Im}w_t|^{-\frac{1}{2}}$ with high probability.

\item This validates that $\tau_{\text{stop},1} = 1$ with high probability, which in turn implies that $\max_{a,b}|E_t(z)_{ab}| \prec W^{-\frac{7}{4}}|\text{Im}w_1|^{-\frac{3}{2}}$ (Theorem 2).
\end{enumerate}

The key insight is that the error bound from stopping time $\tau_{\text{stop},1}$ feeds into the estimates for $\tau_{\text{stop},2}$, allowing us to control $\max_{a,b}|G_t(z)_{ab} - \delta_{ab}m(z)|^2$ (Theorem 4). This demonstrates how the carefully chosen stopping times create a robust framework for establishing quantum diffusion and delocalization results.

The assumption $W \gg N^{8/11}$ arises from the need to ensure that:

\begin{align}
W^{-\frac{7}{4}}\eta^{-\frac{3}{2}} \ll W^{-1}\eta^{-\frac{1}{2}}
\end{align}

With $\eta \approx W^2N^{-2}$ (corresponding to the diffusion time scale), this inequality becomes:

\begin{align}
W^{-\frac{7}{4}}(W^2N^{-2})^{-\frac{3}{2}} \ll W^{-1}(W^2N^{-2})^{-\frac{1}{2}}\\
\Rightarrow W^{-\frac{7}{4}} \cdot W^{-3}N^{3} \ll W^{-1} \cdot W^{-1}N\\
\Rightarrow W^{-\frac{11}{4}}N^{3} \ll W^{-2}N\\
\Rightarrow W^{-\frac{3}{4}} \ll N^{-2}\\
\Rightarrow W \gg N^{\frac{8}{11}}
\end{align}

This demonstrates how the specific exponent $\frac{8}{11}$ arises naturally from the mathematical structure of the problem.

\newpage
\section{Quantum Diffusion and Delocalization}
\label{sec:deloc}
